---
title: "Statistics: Two sample t-test example"
author: "Scott Schumacker"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This project will show an example two sample t-test analyzing blood pressures from two example populations: People under 50 years of age and people above 50 years of age.

***Loading Packages***
```{r}
library(ggplot2)
library(dplyr)
library(gridExtra)
```

## Creating a data set

We can start by creating two sample data sets. We can use the rnorm function to create a normal distribution around two different sample means. In this example, we can keep things simple and keep both sample groups to the same sample size of 100 people. In addition, for this example we will only be analyzing the difference in diastolic blood pressure (lower number).
```{r cars}
under50 <- rnorm(100, mean = 80, sd = 1)
under50 <- as.data.frame(under50)
above50 <- rnorm(100, mean = 90, sd = 1)
above50 <- as.data.frame(above50)
```

After obtaining our data set we can visually inspect the distributions to better understand the shape of the data.
```{r}
P1 <- ggplot(under50, aes(x = under50)) +
  geom_histogram(color = "black", fill = "#92C9B1") +
  theme_classic() +
  xlab("Diastolic Blood Pressure") +
  ggtitle("Under 50 Years Old")

P2 <- ggplot(above50, aes(x = above50)) +
  geom_histogram(color = "black", fill = "#4F759B") +
  theme_classic() +
  xlab("Diastolic Blood Pressure") +
  ggtitle("Above 50 Years Old")

grid.arrange(P1, P2)
```

After visualizing the data we can see that both distributions follow a normal distribution pattern. We can confirm that both groups follow a normal distribution by testing for normality and running the Shaprio-Wilk Test.

### Checking for normality and homoscedasticity

***Checking for normality***
```{r}
shapiro.test(under50$under50)
shapiro.test(above50$above50)
```

Because our p-value is above our alpha of 0.05, we can assume that the data follows a normal distribution.

Now that we know that both sample groups follow a normal distribution, we can check the assumption of homogeneity of variance (homoscedasticity) before running our t-test. To test homogeneity of variance, we can run Fisher's F-Test.

***Checking for homoscedasticity***
```{r}
var.test(under50$under50, above50$above50)
```

Because our p-value is above our alpha of 0.05, we can assume homogeneity of variance between the two groups.

Now that we know that our data is randomly sampled, follows a normal distribution, is continuous, and there is homogeneity of variance between the two groups, we can run the two sample t-test.

### Statistical Analysis

Before running the t-test we need to specify our hypothesis:
H0: The difference in the means between the groups is 0 (null hypothesis)
Ha: The difference in the means between the groups is not 0 (alternative hypothesis)

We can set our alpha to 0.05 for this example. This means that we are looking for a p-value below 0.05 to reject the null hypothesis.

***Student's t-test***
```{r}
testOut <- t.test(under50$under50, above50$above50, var.equal = TRUE)
testOut
```

### Results

Our p-value is less than 0.01 which means that we can reject our null hypothesis. We can assume that there is a statistically significant difference between the two groups in terms of blood pressure. 